\documentclass[11pt,a4paper]{article}

\usepackage[margin=1in]{geometry}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{microtype}
\usepackage{hyperref}
\usepackage{amsmath,amssymb,amsfonts,amsthm}
\usepackage{mathtools}
\usepackage[nameinlink,noabbrev]{cleveref}
\usepackage{booktabs}
\usepackage{array}
\usepackage{longtable}
\usepackage{enumitem}
\usepackage{siunitx}
\usepackage{listings}
\usepackage{xcolor}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{float}
\usepackage{parskip}
\usepackage{tikz}
\usetikzlibrary{shapes.geometric, arrows}

\hypersetup{
    colorlinks=true,
    linkcolor=blue,
    urlcolor=cyan,
    citecolor=blue,
    pdftitle={InterLink: A Zero-Knowledge Atomic Cross-Chain Interoperability Protocol with Unified Liquidity and Incentive-Aligned Tokenomics},
    pdfauthor={MeridianAlgo},
    pdfsubject={Blockchain Interoperability Research Paper},
    pdflang={en}
}

\definecolor{codegreen}{rgb}{0,0.6,0}
\definecolor{codegray}{rgb}{0.5,0.5,0.5}
\definecolor{codepurple}{rgb}{0.58,0,0.82}
\definecolor{backcolour}{rgb}{0.95,0.95,0.92}

% Define Rust language for listings
\lstdefinelanguage{Rust}{
  morekeywords={abstract,as,async,await,become,box,break,const,continue,crate,do,dyn,else,enum,extern,false,fn,for,if,impl,in,let,loop,macro,match,mod,move,mut,pub,ref,return,self,Self,static,struct,super,trait,true,type,typeof,unsafe,unsized,use,virtual,where,while,yield},
  sensitive=true,
  morecomment=[l]{//},
  morecomment=[s]{/*}{*/},
  morestring=[b]",
  morestring=[b]',
}
\lstdefinestyle{mystyle}{
    backgroundcolor=\color{backcolour},
    commentstyle=\color{codegreen},
    keywordstyle=\color{magenta},
    numberstyle=\tiny\color{codegray},
    stringstyle=\color{codepurple},
    basicstyle=\ttfamily\footnotesize,
    breaklines=true,
    captionpos=b,
    keepspaces=true,
    numbers=left,
    numbersep=5pt,
    showstringspaces=false,
    tabsize=2,
    language=Rust,
    frame=single,
    framesep=5pt,
    framerule=0.5pt
}
\lstset{style=mystyle}

\setlength{\parindent}{0pt}
\setlength{\parskip}{8pt plus 2pt minus 1pt}

\title{
    \vspace{2in}
    \textbf{\Huge InterLink Protocol}\\
    \vspace{0.5in}
    \large A Zero-Knowledge Atomic Interoperability Framework\\
    \vspace{0.2in}
    \textit{Trustless. Scalable. Unified.}\\
    \vspace{1in}
    \textbf{MeridianAlgo Research Lab}\\
    \small Version 6.0 --- Technical Compendium --- February 2026
}
\author{}
\date{}

\begin{document}

\maketitle
\begin{titlepage}
\maketitle
\end{titlepage}
\newpage

\tableofcontents
\newpage

\phantomsection
\addcontentsline{toc}{section}{Abstract}
\begin{abstract}
\section*{Abstract}
The rapid proliferation of blockchain protocols has catalyzed a decentralized revolution, enabling trustless value transfer and immutable computation. However, this cambrian explosion of Layer-1 (L1) and Layer-2 (L2) networks has resulted in severe fragmentation, isolating capital, users, and state within disparate execution environments. The prevailing "multi-chain" thesis has inadvertently created a "siloed-chain" reality, where liquidity is fractured, and user experience is marred by complex, high-latency bridging mechanisms. Conventional interoperability solutions, relying heavily on multisig federation or optimistic verification, introduce systemic latency and critical centralization vectors. History has shown these "trusted" bridges to be the weakest link in the crypto-economic stack, accounting for billions of dollars in lost value due to key compromises and smart contract vulnerabilities.

This paper presents \textbf{InterLink (ILINK)}, a hyper-structure for cross-chain composability powered by \textbf{recursive zero-knowledge proofs (zk-SNARKs)}. By utilizing the \textbf{halo2} proving system, InterLink enables trustless, atomic state transitions between heterogeneous ledgers without relying on honest majority assumptions for verification. Our architecture introduces a decentralized, stake-weighted \textbf{Relayer Network} incentivized by a novel, deflationary token model, coupled with a high-performance \textbf{Solana Execution Hub} for global state management. 

We rigorously define the cryptographic primitives, circuit constraints, and economic game theory underpinning the protocol. Specifically, we demonstrate how InterLink achieves O(1) verification cost for batched cross-chain intents while preserving complete sovereign security. This system does not merely act as a bridge; it functions as the foundational substrate for the Unified Liquidity Layer of the decentralized web, allowing developers to build "omnichain" applications that are agnostic to the underlying chain. The results presented herein establish a new standard for interoperability: one that is mathematically verifiable, economically robust, and scalable to billions of users.

\vspace{0.5in}
\textbf{Keywords:} Zero-Knowledge Proofs, Recursive SNARKs, Atomic Composability, Rust Systems Engineering, Solana Anchor, Liquidity Unification, Hyper-Deflationary Tokenomics, Game-Theoretic Security, Blockchain Interoperability, Decentralized Finance (DeFi).
\end{abstract}

\newpage

\section{Introduction}
\label{sec:intro}

The digital landscape is undergoing a fundamental transformation, shifting from centralized, opaque databases to decentralized, transparent ledgers. This shift, powered by blockchain technology, promises to redefine how value is stored, transferred, and programmed. However, as the ecosystem matures, it faces critical bottlenecks that threaten to stifle its potential. To understand the necessity of the \textbf{InterLink Protocol}, one must first understand the underlying technology it seeks to unite: the blockchain itself.

\subsection{The Blockchain Paradigm: A Detailed Overview}
At its core, a blockchain is a distributed, immutable ledger that records transactions across a network of computers. Unlike traditional databases controlled by a central authority (like a bank or a government), a blockchain is maintained by a decentralized network of nodes. Each node stores a copy of the entire ledger, ensuring that no single entity can manipulate the data.

\subsubsection{Cryptographic Foundations}
The integrity of a blockchain relies on two primary cryptographic pillars: hashing and public-key cryptography. 
\begin{itemize}
    \item \textbf{Hashing:} A hash function takes an input of any length and produces a fixed-size string of characters (the hash). This process is deterministic (same input always yields same output) but irreversible (cannot derive input from output). In a blockchain, each block contains a cryptographic hash of the previous block, creating a "chain." If a malicious actor attempts to alter a transaction in a past block, the hash of that block would change, invalidating all subsequent blocks and alerting the network to the tampering.
    \item \textbf{Public-Key Cryptography:} Users interact with the blockchain using pairs of cryptographic keys: a public key (their address) and a private key (their password). Transactions are signed with the private key, proving ownership of the funds without revealing the key itself. This ensures that only the rightful owner can authorize a transfer.
\end{itemize}

\subsubsection{Consensus Mechanisms}
To agree on the state of the ledger without a central authority, blockchains use consensus algorithms.
\begin{itemize}
    \item \textbf{Proof of Work (PoW):} Pioneered by Bitcoin, PoW requires miners to solve complex mathematical puzzles to validate transactions and add new blocks. This process is energy-intensive but extremely secure, as attacking the network requires overwhelming computational power.
    \item \textbf{Proof of Stake (PoS):} Used by Ethereum and Solana, PoS requires validators to "stake" (lock up) their tokens as collateral. Validators are chosen to create blocks based on the amount they stake. Malicious behavior results in the loss of staked tokens (slashing). PoS is more energy-efficient and scalable than PoW.
\end{itemize}

\subsubsection{Smart Contracts and Turing Completeness}
While Bitcoin introduced the concept of decentralized money, Ethereum expanded it to decentralized computation. Smart contracts are self-executing programs stored on the blockchain that run when predetermined conditions are met. They enable complex logic, such as decentralized exchanges (DEXs), lending protocols, and governance systems, to operate autonomously. This "Turing Completeness" allows developers to build any application imaginable on top of the blockchain, leading to the explosion of Decentralized Finance (DeFi) and Non-Fungible Tokens (NFTs).

\subsection{The Fragmentation Crisis: The Problems of Success}
The success of blockchain technology has paradoxically created its biggest challenge: fragmentation. As developers sought to overcome the scalability limits of early blockchains (like Ethereum's low transaction throughput and high fees), they launched dozens of alternative Layer-1 blockchains (Solana, Avalanche, Cosmos) and Layer-2 scaling solutions (Optimism, Arbitrum, zkSync).

\subsubsection{The Siloed State Problem}
Each blockchain operates as a sovereign digital nation with its own rules, consensus mechanism, and state. They are largely unaware of each other. An asset on Ethereum cannot natively exist on Solana. To move value, users must rely on "bridges" that lock assets on one chain and mint a synthetic representation on another. This creates "siloed liquidity," where capital is trapped in isolated pools, reducing efficiency and increasing slippage for traders.

\subsubsection{The Security Vulnerability of Bridges}
Traditional bridges have proven to be the Achilles' heel of the crypto ecosystem. Most rely on a "middleman" approach: a set of trusted validators (often a small multisig of 5-10 people) who verify transactions. If these validators are compromised or collude, they can steal all the funds locked in the bridge. This centralization risk contradicts the core ethos of blockchain. In 2022 alone, bridge hacks accounted for over \$2 billion in losses, highlighting the desperate need for a trustless solution.

\subsection{The Genesis of InterLink: A Proposal for Unification}
We propose \textbf{InterLink} as a robust solution to the interoperability trilemma—Scalability, Security, and Generalizability—without compromise. Recognizing that the future is not "multi-chain" (many isolated chains) but "cross-chain native" (one unified network of chains), we are currently architecting this protocol to serve as the connective tissue of the decentralized web.

\subsubsection{From Trust to Truth}
Existing solutions ask users to trust a third party. InterLink asks users to trust mathematics. By leveraging \textbf{Zero-Knowledge Proofs (zk-SNARKs)}, specifically the \textbf{Halo2} proving system, InterLink is designed to allow one blockchain to cryptographically verify the state of another. In our proposed architecture, a relayer generates a succinct proof that a transaction occurred on Ethereum, and the Solana blockchain verifies this proof in milliseconds. This removes the need for trusted validators. If the proof is valid, the transaction is valid. Period.

\subsubsection{Unified Liquidity}
InterLink is not envisioned just as a message-passing protocol; it is being built as a liquidity unification layer. By using Solana as a high-performance execution hub, we aim to aggregate liquidity from all connected chains into a single, efficient market. A user will be able to swap a token on Ethereum for a token on Cosmos in a single atomic transaction, with the trade routed and settled through the InterLink Hub on Solana.

\subsubsection{Sovereignty and Scalability}
We have chosen a modular architecture that respects the sovereignty of each chain while leveraging the speed of Solana. The heavy lifting of proof generation happens off-chain by a decentralized network of relayers, while the verification happens on-chain. This design ensures that InterLink can scale to handle thousands of cross-chain transactions per second without congesting the source or destination chains.

InterLink represents our vision for the "Internet of Blockchains"—a seamless, secure, and scalable network where value flows as freely as information. We are actively developing the core circuits and smart contracts described in this paper to turn this vision into reality.

\section{Preliminaries}
\label{sec:prelim}

To provide a rigorous foundation for the InterLink Protocol, we must first establish the mathematical and cryptographic primitives upon which the system is built. This section details the specific choices made regarding elliptic curves, proving systems, and blockchain runtime environments.

\subsection{The Concept of Zero-Knowledge}
Before delving into the technicalities of SNARKs, it is crucial to understand the intuition behind Zero-Knowledge Proofs.

\subsubsection{Intuition: The Ali Baba Cave}
Imagine a cave shaped like a ring with a magic door blocking the path on the far side. The door only opens if you know the secret password.
\begin{itemize}
    \item \textbf{Prover (Peggy):} Wants to prove she knows the password without revealing it.
    \item \textbf{Verifier (Victor):} Wants to verify Peggy knows the password.
\end{itemize}
\textbf{The Protocol:}
1. Peggy enters the cave and randomly chooses path A or path B. Victor waits outside.
2. Victor enters and shouts "Come out path A!" (or B).
3. If Peggy knows the password, she can always comply, opening the door if necessary to cross to the correct path.
4. If Peggy does NOT know the password, she has a 50\% chance of being on the wrong side and getting caught.

By repeating this test $n$ times, the probability of Peggy cheating becomes $2^{-n}$. If $n=40$, the chance is less than one in a trillion. Thus, Victor is convinced Peggy knows the password, but he never learned the password itself (Zero-Knowledge).

\subsection{zk-SNARKs: A Technical Deep Dive}
InterLink utilizes \textbf{zk-SNARKs} (Zero-Knowledge Succinct Non-Interactive Arguments of Knowledge). Let us break down each term:

\begin{description}
    \item[Zero-Knowledge:] As explained above, the verifier learns nothing about the private inputs (witness) other than the fact that the statement is true. In InterLink, this means the Solana blockchain learns that "Transaction X happened on Ethereum" without needing to see the entire Ethereum blockchain history.
    
    \item[Succinct:] The proof size is incredibly small (constant or logarithmic size), and verification time is fast (milliseconds). For example, a proof attesting to the validity of 1,000 transactions might only be a few hundred bytes. This is critical for on-chain verification where storage and compute are expensive.
    
    \item[Non-Interactive:] Unlike the Cave example where Victor shouts challenges, a SNARK allows Peggy to generate a proof once, publish it, and anyone can verify it at any time without further interaction. This is achieved via the Fiat-Shamir heuristic, which turns an interactive protocol into a non-interactive one by using a cryptographic hash function to simulate the verifier's randomness.
    
    \item[Argument:] A slightly weaker notion than a "Proof." A proof holds against a computationally infinite adversary. An argument holds against a "computationally bounded" adversary (one who cannot break encryption in polynomial time). For all practical purposes in blockchain, this is sufficient.
    
    \item[of Knowledge:] The proof guarantees that the prover actually \textit{knows} the witness (the private data), not just that such a witness exists.
\end{description}

\subsection{Elliptic Curve Cryptography (ECC)}
InterLink leverages specific elliptic curves to optimize for both Ethereum compatibility and recursive proof composition.

\subsubsection{BN254 (Alt-Bn128)}
For interactions with the Ethereum Virtual Machine (EVM), we utilize the BN254 curve. This is a pairing-friendly curve defined over a finite field $\mathbb{F}_q$. It allows for efficient pairing operations $e: \mathbb{G}_1 \times \mathbb{G}_2 \rightarrow \mathbb{G}_T$, which are essential for verifying KZG commitments on-chain.
$$ y^2 = x^3 + 3 \mod q $$
Ethereum has precompiled contracts (0x06, 0x07, 0x08) specifically optimized for BN254 operations, enabling gas-efficient verification.

\subsubsection{Pallas and Vesta (The Pasta Curves)}
For the recursive proof generation inside the Relayer Network, we employ the Pallas and Vesta curves. These form a cycle of curves: the scalar field of Pallas is the base field of Vesta, and vice versa.
\begin{itemize}
    \item \textbf{Pallas:} $y^2 = x^3 + 5$
    \item \textbf{Vesta:} $y^2 = x^3 + 5$ (over a different field)
\end{itemize}
This cycle property allows us to efficiently verify a proof generated over Pallas inside a circuit defined over Vesta, avoiding the massive overhead of non-native field arithmetic ("wrong-field" arithmetic). This is the cornerstone of our "Halo2" recursion strategy.

\subsection{Polynomial Commitment Schemes (PCS)}
A PCS allows one to commit to a polynomial $P(X)$ and later prove that $P(z) = y$.

\subsubsection{KZG (Kate-Zaverucha-Goldberg)}
Used for the final "outer" proof submitted to Ethereum.
\begin{itemize}
    \item \textbf{Pros:} Constant size proofs (48 bytes), constant time verification.
    \item \textbf{Cons:} Requires a Trusted Setup (Structured Reference String).
\end{itemize}

\subsubsection{IPA (Inner Product Argument)}
Used for the internal "inner" proofs within the recursive tree.
\begin{itemize}
    \item \textbf{Pros:} Transparent (no trusted setup), relies only on discrete log hardness.
    \item \textbf{Cons:} $O(\log n)$ verification time (slower than KZG).
\end{itemize}
InterLink uses a hybrid approach: IPA for recursion (accumulation) and a final wrap into KZG/Groth16 for cheap on-chain verification.

\subsection{Solana Runtime (Sealevel)}
Solana differs fundamentally from EVM-based chains.
\begin{itemize}
    \item \textbf{Stateless Execution:} Programs (smart contracts) are stateless; state is stored in separate "Accounts."
    \item \textbf{Parallelism:} Transactions specify all accounts they will read or write upfront. This allows the runtime to schedule non-overlapping transactions in parallel.
    \item \textbf{PDAs (Program Derived Addresses):} Addresses deterministically derived from a program ID and seeds (e.g., `hash("user_vault", user_pubkey)`). This allows us to map a user's Ethereum address to a specific Solana account without needing a private key for that account.
\end{itemize}

\section{System Architecture}
\label{sec:arch}

\subsection{The Hub-and-Spoke Topology}
InterLink adopts a Hub-and-Spoke network topology to minimize complexity.
\begin{itemize}
    \item \textbf{Hub:} The Solana Blockchain. It acts as the global clearinghouse for state. All cross-chain messages are routed through the Hub.
    \item \textbf{Spokes:} Connected blockchains (Ethereum, Arbitrum, Optimism, Cosmos Hub, Sui).
\end{itemize}
This $O(N)$ connection complexity is superior to the $O(N^2)$ complexity of pairwise bridges. Adding a new chain only requires establishing a connection to the Solana Hub.

\subsection{Detailed Component Breakdown}

\subsubsection{Source Chain Gateway (The Spoke)}
Deployed on every supported chain (e.g., Ethereum), the Gateway is a set of smart contracts responsible for:
\begin{itemize}
    \item \textbf{Asset Custody:} Holding user funds in a non-custodial vault.
    \item \textbf{Event Emission:} Emitting canonical logs (e.g., `LogMessagePublished`) that Relayers can subscribe to.
    \item \textbf{Message Ingestion:} Receiving verified instructions from the Hub to unlock assets.
\end{itemize}
\textbf{Safety Module:} The Gateway includes a "Guardian" role (controlled by the DAO) that can pause the contract in the event of a detected anomaly, providing an emergency circuit breaker.

\subsubsection{The Relayer Network (The Transport)}
A decentralized network of off-chain nodes.
\begin{itemize}
    \item \textbf{Listener:} Uses WebSocket connections to full nodes of connected chains to listen for Gateway events.
    \item \textbf{Prover (The Engine):} The most computationally intensive component. It runs the Halo2 proving stack.
        \begin{itemize}
            \item \textit{Input:} Block header, Merkle branch, Transaction content.
            \item \textit{Process:} Generates a ZK-proof attesting to the validity of the transaction and its inclusion in the canonical chain.
            \item \textit{Recursion:} Multiple transaction proofs are aggregated into a single "Batch Proof" to amortize gas costs.
        \end{itemize}
    \item \textbf{Submitter:} Monitors the Solana Hub. When a batch is ready, it submits the proof and the public inputs as a Solana transaction.
\end{itemize}

\subsubsection{3. The Solana Execution Hub (The Core)}
Implemented as an Anchor Program on Solana.
\begin{itemize}
    \item \textbf{Verifier Contract:} Contains the verification keys for the supported circuits. It executes the pairing checks (or validates the accumulation scheme).
    \item \textbf{State Registry:} Maintains a mapping of `sequence_number -> message_hash` to prevent replay attacks.
    \item \textbf{Liquidity Pools:} Implements a concentrated liquidity AMM (CLMM) where cross-chain swaps are executed.
    \item \textbf{Messenger:} Queues outbound messages to destination chains.
\end{itemize}

\subsection{Cross-Chain Message Passing Lifecycle}
To fully understand the system, we must examine the lifecycle of a message as it traverses the stack.

\begin{enumerate}
    \item \textbf{Inbound Trigger:} A user on Arbitrum calls `InterLinkRouter.send(dest=Solana, payload=...)`. The contract emits an event with a unique sequence number.
    \item \textbf{Block Finality:} Relayers wait for the Arbitrum sequencer to post the batch to Ethereum L1, and for that L1 block to be finalized (approx. 12-15 mins for full security, or soft-finality for optimistic mode).
    \item \textbf{Witness Extraction:} The Relayer queries an Archive Node for the Merkle Proof of the transaction receipt against the block's `receiptsRoot`.
    \item \textbf{Proof Composition:}
        \begin{itemize}
            \item \textit{Circuit 1 (Transaction):} Verifies the Merkle Path.
            \item \textit{Circuit 2 (Consensus):} Verifies the Block Header signatures.
            \item \textit{Circuit 3 (Aggregation):} Compresses Circuit 1 and Circuit 2 into a single proof.
        \end{itemize}
    \item \textbf{Solana Submission:} The final proof is submitted to the InterLink Hub. The transaction includes the `public_inputs` (the message hash, the block root).
    \item \textbf{On-Chain Verification:} The Solana Runtime executes the `verify` instruction. If successful, the message hash is marked as "verified" in the PDA state.
    \item \textbf{Execution:} An "Executor" bot (can be the same as Relayer) sees the verified state and executes the payload (e.g., minting a token, executing a swap).
\end{enumerate}

\subsection{Consensus Verification in Circuits}
A unique feature of InterLink is "Light Client in a Circuit." Instead of just verifying a transaction signature, we verify the consensus of the source chain.
\begin{itemize}
    \item \textbf{Ethereum (Sync Committee):} The circuit verifies the BLS signatures of the current Sync Committee (512 validators) to confirm the block header validity.
    \item \textbf{Cosmos (Tendermint):} The circuit verifies the Ed25519 signatures of the validator set (voting power > 2/3).
\end{itemize}
This means InterLink inherits the full security of the connected chain's consensus, rather than relying on a small multisig.

\section{Rust Implementation \& Engineering}
\label{sec:impl}

The engineering philosophy of InterLink prioritizes safety, performance, and maintainability. We leverage the Rust programming language across the entire stack—from the off-chain Relayers to the on-chain Solana programs—to ensure memory safety and type correctness.

\subsection{Circuit Engineering with Halo2}
The heart of InterLink is the ZK-circuit. We use the \textbf{halo2-ce} (Community Edition) crate for its flexibility and support for the KZG commitment scheme.

\subsubsection{Circuit Configuration}
A Plonk-ish circuit is defined by a matrix of columns and rows. We define a custom `InterlinkConfig` struct to manage these resources.
\begin{lstlisting}[language=Rust, caption={Halo2 Circuit Configuration}]
#[derive(Clone, Debug)]
pub struct InterlinkConfig {
    // Advice columns (Witnesses): Private inputs like secret keys or Merkle paths
    pub advice: [Column<Advice>; 5],
    // Instance columns (Public Inputs): Root hash, Nullifier, etc.
    pub instance: Column<Instance>,
    // Fixed columns (Constants): Selector bits, lookup tables
    pub fixed: [Column<Fixed>; 2],
    // Selectors to toggle gates
    pub s_hash: Selector,
    pub s_verify: Selector,
}

impl InterlinkConfig {
    pub fn configure(meta: &mut ConstraintSystem<F>) -> Self {
        let advice = [0; 5].map(|_| meta.advice_column());
        let instance = meta.instance_column();
        let fixed = [0; 2].map(|_| meta.fixed_column());
        let s_hash = meta.selector();

        meta.enable_equality(instance);
        for c in &advice { meta.enable_equality(*c); }

        // Define a Custom Gate: Poseidon Hash Round
        meta.create_gate("poseidon_round", |meta| {
            let s = meta.query_selector(s_hash);
            let a = meta.query_advice(advice[0], Rotation::cur());
            let b = meta.query_advice(advice[1], Rotation::cur());
            let out = meta.query_advice(advice[2], Rotation::cur());
            
            // Constraint: out = (a + b)^5 (simplified S-box)
            vec![s * (out - (a + b).pow([5]))]
        });

        Self { advice, instance, fixed, s_hash, s_verify: meta.selector() }
    }
}
\end{lstlisting}

\subsubsection{The Merkle Inclusion Chip}
We encapsulate logic into "Chips." The `MerkleChip` is responsible for verifying that a leaf exists in a tree. It takes a `Path` (witness) and a `Root` (public input) and enforces the hashing path.
\begin{itemize}
    \item \textbf{Input:} Leaf Node $L$, Path Elements $P_0, ..., P_n$, Path Indices $I_0, ..., I_n$ (0 for left, 1 for right).
    \item \textbf{Logic:} For each level, compute $H_{i+1} = \text{Hash}(L_i, P_i)$ if $I_i=0$, else $\text{Hash}(P_i, L_i)$.
    \item \textbf{Output:} The computed Root $R_{calc}$.
    \item \textbf{Constraint:} $R_{calc} == R_{public}$.
\end{itemize}

\subsection{Relayer Infrastructure}
The Relayer node is a high-performance Rust binary designed for 24/7 operation.

\subsubsection{Concurrency Model}
We utilize `tokio` for asynchronous runtime execution. The architecture follows a producer-consumer pattern:
\begin{enumerate}
    \item \textbf{Event Watcher (Producer):} Polls Ethereum logs via WebSocket (`ethers-rs`). Pushes events to a bounded `mpsc` channel.
    \item \textbf{Proof Generator (Consumer):} Pulls events, fetches witness data (Merkle proofs) from an archive node, and runs the CPU-intensive proving algorithm. This runs in a separate thread pool (`rayon`) to avoid blocking the async runtime.
    \item \textbf{Transaction Manager:} Queues the generated proof and manages the Solana transaction lifecycle (retries, priority fees).
\end{enumerate}

\begin{lstlisting}[language=Rust, caption={Relayer Proof Generation Logic}]
pub async fn process_event(event: DepositEvent, params: &Params<Eq>) -> Result<Proof> {
    // 1. Fetch Witness Data (Network I/O)
    let merkle_path = eth_client.get_proof(event.address, event.slot).await?;
    
    // 2. Generate Proof (CPU Bound - Offload to thread)
    let proof = tokio::task::spawn_blocking(move || {
        let circuit = InterlinkCircuit::new(event, merkle_path);
        let mut transcript = Blake2bWrite::<_, _, Challenge255<_>>::init(vec![]);
        create_proof(params, &pk, &[circuit], &[&[&public_inputs]], &mut transcript)?;
        Ok(transcript.finalize())
    }).await??;

    Ok(proof)
}
\end{lstlisting}

\subsection{Solana Anchor Development}
The Verification Layer is built with the Anchor framework, which provides safety checks and serialization for Solana programs.

\subsubsection{Instruction Context & CPIs}
When a proof is verified, the program must mint tokens. This requires a Cross-Program Invocation (CPI) to the SPL Token Program.
\begin{lstlisting}[language=Rust, caption={Anchor CPI for Minting}]
#[derive(Accounts)]
pub struct VerifyAndMint<'info> {
    #[account(mut)]
    pub payer: Signer<'info>,
    #[account(mut, seeds = [b"mint_authority"], bump)]
    pub mint_authority: SystemAccount<'info>,
    #[account(mut)]
    pub recipient: Account<'info, TokenAccount>,
    #[account(mut)]
    pub token_mint: Account<'info, Mint>,
    pub token_program: Program<'info, Token>,
}

pub fn handler(ctx: Context<VerifyAndMint>, amount: u64) -> Result<()> {
    // ... Verification Logic ...

    // Mint Tokens via CPI
    let seeds = &[b"mint_authority".as_ref(), &[ctx.bumps.mint_authority]];
    let signer = &[&seeds[..]];

    let cpi_ctx = CpiContext::new_with_signer(
        ctx.accounts.token_program.to_account_info(),
        token::MintTo {
            mint: ctx.accounts.token_mint.to_account_info(),
            to: ctx.accounts.recipient.to_account_info(),
            authority: ctx.accounts.mint_authority.to_account_info(),
        },
        signer,
    );
    token::mint_to(cpi_ctx, amount)?;
    Ok(())
}
\end{lstlisting}

\subsubsection{Compute Budget Optimization}
Solana has a strict compute unit (CU) limit (200k per instruction, 1.4M per transaction). Verifying a Pairing Check is expensive.
\begin{itemize}
    \item \textbf{Strategy:} We use the `ed25519_program` for signature verification where possible, and highly optimized BPF assembly for the pairing engine.
    \item \textbf{Address Lookup Tables (LUTs):} We use LUTs to load more accounts in a single transaction, allowing us to split the verification into multiple instructions if necessary.
\end{itemize}

\subsection{Performance Optimization}
To achieve the sub-second latency required for a high-frequency bridge, we implement several optimization strategies:

\begin{itemize}
    \item \textbf{Parallel Proof Generation:} The Halo2 prover supports multi-threading for FFTs (Fast Fourier Transforms) and MSMs (Multi-Scalar Multiplications). We use `rayon` to saturate all CPU cores during the proving phase.
    \item \textbf{Batching:} Instead of generating one proof per transaction, we aggregate $N$ transactions into a single batch. The verification cost on Solana is paid once per batch, amortizing the cost to $Cost / N$.
    \item \textbf{Pipeline Architecture:} The Relayer uses a pipelined approach (Fetch -> Prove -> Submit). While Batch $i$ is being submitted, Batch $i+1$ is being proven, and Batch $i+2$ is being fetched.
\end{itemize}

\section{Tokenomics: The ILINK Standard}
\label{sec:tokenomics}

The ILINK token is the economic fuel of the InterLink ecosystem. It is designed to capture value from cross-chain activity while securing the network through stake-weighted incentives.

\subsection{Supply Dynamics}
\begin{itemize}
    \item \textbf{Total Supply:} 1,000,000,000 (1 Billion) ILINK.
    \item \textbf{Emission Schedule:} Halving every 2 years, similar to Bitcoin, ensuring long-term scarcity.
    \item \textbf{Distribution:}
        \begin{itemize}
            \item 20\% Team \& Advisors (4-year vesting, 1-year cliff).
            \item 15\% Investors (3-year vesting).
            \item 30\% Ecosystem Growth (Grants, Liquidity Mining).
            \item 35\% Community Treasury (DAO controlled).
        \end{itemize}
\end{itemize}

\subsection{Utility A: The Security Bond}
Relayers must stake ILINK to participate in the network.
$$ \text{Stake}_{min} = 100,000 \text{ ILINK} $$
If a Relayer submits an invalid proof (which is caught by the Verifier), their stake is slashed.
$$ \text{SlashAmount} = \text{Stake} \times 50\% $$
The slashed funds are burned, instantly reducing the total supply. This creates a strong economic disincentive for malicious behavior.

\subsection{Utility B: Gas Abstraction \& The Buy-Back-and-Burn}
Users do not need to hold ILINK to use the bridge. They pay gas in the source chain's native token (e.g., ETH).
\begin{enumerate}
    \item User pays 0.01 ETH fee.
    \item Protocol swaps 0.01 ETH for ILINK on a DEX (e.g., Uniswap, Orca).
    \item **40\% of this ILINK is burned.**
    \item 60\% is distributed to Relayers as a reward for their work.
\end{enumerate}
This mechanism ensures that **Increased Usage $\rightarrow$ Increased Buy Pressure $\rightarrow$ Decreased Supply**.

\subsection{Mathematical Model of Deflation}
Let $V$ be the daily volume of cross-chain transfers.
Let $f$ be the fee rate (e.g., 0.1\%).
Let $P$ be the price of ILINK.
The daily amount of tokens burned, $B$, is:
$$ B = \frac{V \times f \times 0.40}{P} $$
If volume $V$ grows and price $P$ remains constant, the burn rate accelerates. If $P$ increases, the burn rate (in token terms) slows, stabilizing the economy. This is a self-regulating feedback loop.

\subsection{Governance}
ILINK holders govern the protocol via the InterLink DAO.
\begin{itemize}
    \item \textbf{Parameter Tweaking:} Adjusting the fee rate $f$ or the burn percentage.
    \item \textbf{Chain Support:} Voting to deploy Gateways to new chains.
    \item \textbf{Treasury Management:} Allocating funds for audits, bug bounties, and developer grants.
\end{itemize}
Governance uses a quadratic voting mechanism to prevent plutocracy (whale dominance), where the cost of votes is the square of the number of votes: $\text{Cost} = (\text{Votes})^2$.

\section{Security Analysis}
\label{sec:security}

Security is the paramount requirement for any interoperability protocol. InterLink approaches security through a layered model, combining cryptographic hardness with game-theoretic incentives.

\subsection{Game Theoretic Analysis}
We analyze the protocol as a game between a Relayer $R$ and the Protocol $P$.
\begin{itemize}
    \item \textbf{Strategy Space:} $R$ can choose to be $\{$Honest, Malicious$\}$.
    \item \textbf{Payoff Matrix:}
\end{itemize}

\begin{table}[H]
\centering
\begin{tabular}{c|c|c}
 & \textbf{Valid Proof} & \textbf{Invalid Proof} \\ \hline
\textbf{Accept} & $R$: +Fee, $P$: +UserTrust & $R$: +Vault, $P$: -Collapse \\ \hline
\textbf{Reject} & $R$: -Gas, $P$: -Liveness & $R$: -Stake, $P$: +Security \\ \hline
\end{tabular}
\caption{Relayer-Protocol Payoff Matrix}
\end{table}

Let $S$ be the staked amount. Let $G$ be the potential gain from a hack (Total Value Locked). Let $\mathbb{P}_{verif}$ be the probability the verifier catches a cheat. In a ZK system, $\mathbb{P}_{verif} = 1$ (soundness).
Therefore, the expected payoff for malicious behavior is:
$$ E[Malicious] = (1 - \mathbb{P}_{verif}) \times G - \mathbb{P}_{verif} \times S $$
Since $\mathbb{P}_{verif} \approx 1$, $E[Malicious] \approx -S$.
A rational actor will never attack the system as long as the cost of generating a false proof (breaking the cryptography) is sufficiently high (computationally infeasible).

\subsection{Specific Attack Vectors & Mitigations}

\subsubsection{1. Long-Range Attacks}
\textbf{Threat:} An attacker creates a private fork of the source chain starting from a point far in the past, eventually overtaking the canonical chain with a different state.
\textbf{Mitigation:} InterLink implements \textbf{Checkpointing}. The Solana Hub stores a history of finalized block headers. Once a header is finalized on Solana, it cannot be reverted, even if the source chain reorgs deep in the past (beyond the finality threshold).

\subsubsection{2. Data Availability (DA) Attacks}
\textbf{Threat:} A Relayer provides a valid proof of a state transition, but the data required to reconstruct that state is withheld from the public.
\textbf{Mitigation:} InterLink requires \textbf{DA Attestation}. The proof must include a commitment to the data being available on a robust DA layer (like Ethereum CallData or Celestia). The circuit verifies that the data was posted.

\subsubsection{3. Censorship}
\textbf{Threat:} Relayers collude to ignore transactions from specific users.
\textbf{Mitigation:} \textbf{Open Relayer Set.} Anyone can become a Relayer by staking ILINK. If the major Relayers censor a user, a new Relayer can spin up, process that specific transaction, collect the fee, and shut down. This "Free Market" approach guarantees liveness.

\subsection{Auditing & Formal Verification}
To mitigate Smart Contract Risk, InterLink undergoes a rigorous pipeline:
\begin{enumerate}
    \item \textbf{Static Analysis:} Automated scanning with tools like Slither (for Solidity) and Soteria (for Solana).
    \item \textbf{Fuzzing:} Property-based testing using `proptest` and `trident` to feed random inputs to the contracts and circuits, checking for invariants (e.g., "Total supply never exceeds 1B").
    \item \textbf{Formal Verification:} We are working with \textit{Veridise} to mathematically prove that our circuits exactly match the intended logic (constraint satisfaction).
\end{enumerate}

\section{Comparison with State of the Art}
\label{sec:compare}

InterLink competes in a crowded market. However, its architecture offers distinct advantages over incumbents.

\begin{table}[H]
\centering
\small
\begin{tabular}{@{}lcccccc@{}}
\toprule
\textbf{Protocol} & \textbf{Trust Model} & \textbf{Latency} & \textbf{Gas Cost} & \textbf{Finality} & \textbf{Setup} & \textbf{Extensibility} \\ \midrule
\textbf{InterLink} & \textbf{ZK-Light Client} & \textbf{Medium} & \textbf{Low (O(1))} & \textbf{Src Finality} & \textbf{Trustless} & \textbf{High} \\
LayerZero v2 & DVNs (Multi-sig) & Low & Low & Instant & Permissioned & High \\
Axelar & PoS Validators & Medium & Medium & 1-2 mins & Trusted & Medium \\
Wormhole & Guardian Set (19/19) & Low & Low & Instant & Trusted & High \\
Cosmos IBC & Light Client & Low & Low & Instant & Trustless & Low \\ \bottomrule
\end{tabular}
\caption{Comprehensive Competitive Analysis}
\label{tab:comparison_detailed}
\end{table}

\subsection{vs. LayerZero}
LayerZero uses "Ultra Light Nodes" (ULNs). In v1, security relied on the independence of the Oracle and Relayer. In v2, it relies on Decentralized Verifier Networks (DVNs). While flexible, it ultimately trusts a set of external parties (like Google Cloud or Polyhedra). InterLink \textit{replaces} these parties with a ZK proof. You don't trust the Relayer; you verify their work.

\subsection{vs. Axelar}
Axelar is a blockchain itself. It uses a PoS validator set to vote on state changes. This is secure but adds an intermediary consensus layer. If the Axelar chain halts or is taken over, the bridge stops. InterLink is not a chain; it is a protocol. Its security inherits directly from the source and destination chains.

\subsection{vs. Wormhole}
Wormhole relies on a "Guardian Set" of 19 reputable entities. While they have a strong track record, it is a Proof-of-Authority system. It is not censorship-resistant by design. InterLink's open relayer network is permissionless.

\subsection{vs. Cosmos IBC}
IBC is the gold standard for trustless interoperability. However, it is expensive to verify on non-Cosmos chains (like Ethereum) due to the cost of checking Ed25519 signatures in EVM. InterLink is effectively "IBC with ZK." We use ZK proofs to compress the consensus verification, making it affordable to run an IBC-style light client on any chain.

\section{Conclusion}

InterLink represents the maturity of the interoperability landscape. We have moved from the "Wild West" of multisig bridges, plagued by hacks and centralization, to the era of \textbf{Verifiable Interoperability}.

By combining the cryptographic guarantees of \textbf{halo2 zk-SNARKs} with the high-performance execution environment of \textbf{Solana}, InterLink solves the Trilemma. It offers the trustlessness of a light client with the cost-effectiveness of a multisig. The result is a Unified Liquidity Layer that empowers developers to build applications that transcend chain boundaries.

The future of blockchain is not about which chain wins. It is about how they all work together. InterLink is the common language that allows them to speak.

\subsection{Future Work}
Our roadmap includes several key research directions:
\begin{itemize}
    \item \textbf{Privacy-Preserving Bridging:} integrating \textbf{Aztec}-style privacy proofs to allow users to bridge assets without revealing the amount or the recipient.
    \item \textbf{Hardware Acceleration:} Developing FPGA and ASIC designs specifically optimized for the FFT (Fast Fourier Transform) and MSM (Multi-Scalar Multiplication) operations in the Halo2 prover, aiming to reduce proof generation time to sub-second latency.
    \item \textbf{L3 Fractal Scaling:} Allowing app-chains to settle directly on the InterLink Hub, using it as a DA and Settlement layer.
\end{itemize}

\begin{thebibliography}{99}

\bibitem{halo2}
Electric Coin Co. \textit{The Halo2 Proving System}. 2024. \url{https://zcash.github.io/halo2/}

\bibitem{plonk}
Gabizon, A., Williamson, Z., Ciobotaru, O. \textit{PLONK: Permutations over Lagrange-bases for Oecumenical Noninteractive arguments of Knowledge}. 2019.

\bibitem{solana}
Yakovenko, A. \textit{Solana: A new architecture for a high performance blockchain}. 2018.

\bibitem{layerzero}
Zarick, R., et al. \textit{LayerZero: Trustless Omnichain Interoperability Protocol}. 2021.

\bibitem{kzg}
Kate, A., Zaverucha, G.M., Goldberg, I. \textit{Constant-Size Commitments to Polynomials and Their Applications}. 2010.

\bibitem{wormhole}
Wormhole Team. \textit{Wormhole: Generic Message Passing}. 2020.

\end{thebibliography}

\end{document}
